---
title: "P8105_hw2_zf2212"
author: "Catherine"
date: "9/28/2018"
output: html_document
---

## Problem 1
```{r import data}
library(tidyverse)

nyc_data = read_csv(file = "./data/NYC_Transit_Subway_Entrance_And_Exit_Data.csv") %>% 
  janitor::clean_names() %>%
  select(line, station_name, station_latitude, station_longitude, route1:route11, entry, vending, entrance_type, ada, ada_notes) %>% 
  mutate(entry, entry = ifelse(entry == "YES", TRUE, FALSE)) 
```
The dataset above contains information about NYC transit, including line, station name, station latitude, station longitude, routes, entry, vending, entrance type and ADA compliance information. 
I gather the route1 to route11 information. I retain useful data. Convert the entry variable from YES or NO to a logical variable.
The dimension of the dataset is `r dim(nyc_data)`.

```{r distinct station}
distinct_station = distinct(nyc_data, nyc_data$line, nyc_data$station_name, .keep_all = TRUE)  
station_number = nrow(distinct_station)
ada_compliant = sum(distinct_station$ada)  
vending_allow = nrow(filter(distinct_station, entry == TRUE, vending == "NO")) / station_number
```

There are `r station_number` distinct stations. 
Among these stations, `r ada_compliant` are ADA compliant. 
`r vending_allow` of station entrances / exits without vending allow entrance.

```{r A train}
distinct_a = gather(nyc_data, key = route, value = line_name, route1:route11, na.rm = TRUE) %>% 
filter(line_name == "A")
number_a = nrow(distinct_a)
ada_a = nrow(filter(distinct_a, ada == TRUE))
```
There are `r number_a` stations serve the A train. 
Of the stations that serve the A train, there are `r ada_a` are ADA compliant.

## Problem 2
```{r mtw data import}
mtw_data = readxl::read_xlsx("./data/HealthyHarborWaterWheelTotals2017-9-26.xlsx", range = cellranger::cell_cols("Mr. Trash Wheel!A:N")) %>% 
  janitor::clean_names() %>% 
  na.omit(dumpster) %>% 
  mutate(sports_balls, sports_balls = round(sports_balls,0))
```

```{r precipitation data}
# Import data of 2016 and 2017
data_16 = readxl::read_xlsx("./data/HealthyHarborWaterWheelTotals2017-9-26.xlsx", range = "2016 Precipitation!A2:B14") %>% 
  mutate(year = 2016) %>% 
  janitor::clean_names()

data_17 = readxl::read_xlsx("./data/HealthyHarborWaterWheelTotals2017-9-26.xlsx", range = "2017 Precipitation!A2:B14") %>% 
  mutate(year = 2017) %>% 
  janitor::clean_names()
# Combine datasets
data_1617 = rbind(data_16,data_17) %>% 
  mutate(month, month = month.name[month])
```

```{r}

devtools::install_github("p8105/p8105.datasets")
library(p8105.datasets)
data(brfss_smart2010)
```

